---
title: Transposing a dataframe
authors: 
  - admin
date: '2019-10-24'
categories:
  - Quick Solves
tags:
  - Quick Solves
  - purrr
  - transposing
image:
  caption: ""
  focal_point: "Smart"
summary: "Using purrr to transpose a dataframe"
highlight: true
math: true
output:
  blogdown::html_page:
    toc: false
    number_sections: false
    toc_depth: 4
    fig_caption: yes
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_knit$set(eval.after = 'fig.cap', message = FALSE, warning=FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rlang))
```

## The Problem
 
Here is the challenge:

```{r}
data <- data.frame(
  "id" = c(901, 902, 903, "age", "gender", "language"),
  "rater1" = c(7, 9, 9, 21, 1, 1),
  "rater2" = c(9, 9, 9, 39, 2, 2),
  "rater3" = c(9, 9, 9, 38, 2, 1),
  "rater4" = c(9, 9, 9, 33, 2, 1),
  "rater5" = c(2, 9, 9, 21, 2, 1)
)
```

Filter all the ratings with gender 1, or language 1, or gender 1 AND language 1.

## The Solution

The thing is super easy, we need to transpose, the thing transposition is not a valid operation when it comes to data frames, how can we apply this in a data frame using `tidyverse`.

So in order to understand what happens I will run the solution by parts.

```{r}
data %>%
  transpose()
```
The problem using transpose is that the results is a list of lists, so it´s necessary to transform inner list into vectors:

```{r}
data %>% 
  #select raters
  select(contains("rater")) %>%
  #transpose, the problem is that this transform data into lists of lists.
  transpose() %>%
  #unlisting into double vectors
  map(flatten_dbl)
```

Now every list slot can be renamed with the corresponding id:

```{r}
data %>% 
  #select raters
  select(contains("rater")) %>%
  #transpose, the problem is that this transform data into lists of lists.
  transpose() %>%
  #unlisting 
  map(flatten_dbl) %>%
  set_names(data$id)

```
Finally we can reorganize using map_dfc() function that reorder the data into dataframes by column:

```{r}
(
  new_data <- data %>%
    #select raters
    select(contains("rater")) %>%
    #transpose, the problem is that this transform data into lists of lists.
    transpose() %>%
    #unlisting
    map(flatten_dbl) %>%
    set_names(data$id) %>%
    map_dfc( ~ .x)
)
```

Now we can filter accordingly the requested filterings:

```{r}
new_data %>%
  filter(gender == 1)
```
```{r}
new_data %>%
  filter(language == 1)
```

```{r}
new_data %>%
  filter(gender == 1 & language == 1)
```
 
 