---
title: Ugly Untied Dataset
authors: 
  - admin
date: '2019-10-19'
categories:
  - Quick Solves
tags:
  - Quick Solves
  - tidyR
  - left_join()
  - ugly data
image:
  caption: ""
  focal_point: "Smart"
summary: "Using TidyR and dplyr Joins to fix some really ugly Data"
highlight: true
math: true
output:
  blogdown::html_page:
    toc: false
    number_sections: false
    toc_depth: 4
    fig_caption: yes
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_knit$set(eval.after = 'fig.cap', message = FALSE, warning=FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rlang))
```


Even though the Data looks messy and an Intruitive solution didn´t pop up inmediately, It was relatively short to fix.

## The Problem

I want to save some words so I'll go to the source

<center>
```{r echo=FALSE}
blogdown::shortcode('twitter', '1185527244569174017')
```
</center>

So I just replicated the data and did the following:

 * Made data longer eliminating all the NAs that showed up.
 
 * Divided into Intercept and sofa_study_day.
 
 * Joining both together to obtain the 5 records.
 
The code looks like this:

```{r}
data <- tibble::tribble(
  ~Sofa.time.point,     ~Se,   ~Wald, ~p.value,      ~x1,      ~x2,      ~x3,      ~x4,      ~x5,
       "intercept", 0.12395, -24.333,        0,       NA,       NA,       NA, -3.01592,       NA,
       "intercept", 0.13165, -40.045,        0,       NA,       NA,       NA,       NA, -5.27211,
       "intercept", 0.21603,  -7.372,        0,       NA, -1.59253,       NA,       NA,       NA,
       "intercept", 0.23614,  -5.085,        0,       NA,       NA, -1.20082,       NA,       NA,
       "intercept",      NA,      NA,        0,        0,       NA,       NA,       NA,       NA,
  "sofa_study_day", 0.00411, -14.669,        0,       NA,       NA,       NA,       NA, -0.06028,
  "sofa_study_day", 0.00479, -34.798,        0,       NA,       NA,       NA, -0.16685,       NA,
  "sofa_study_day", 0.00615, -39.744,        0, -0.24443,       NA,       NA,       NA,       NA,
  "sofa_study_day", 0.00756,  -9.975,        0,       NA,       NA, -0.07543,       NA,       NA,
  "sofa_study_day", 0.02224, -24.673,        0,       NA,  -0.5488,       NA,       NA,       NA
  )
```

```{r}
tidy_data <- data %>%
  pivot_longer(
    # keeping columns from "Sofa.time.point" to "p.value"
    -(Sofa.time.point:p.value),
    # transform x columns into just one column
    names_to = "x",
    # populate with values
    values_to = "values",
    # dropping NAs
    values_drop_na = TRUE
  )

# "intercept" data
tidy_data %>%
  filter(Sofa.time.point == "intercept") %>%
  left_join(
    #joined with "sofa_study_day"
    tidy_data %>%
      filter(Sofa.time.point == "sofa_study_day"),
    # joining by "x"
    by = "x",
    # adding identifiers to columns having the same name
    suffix = c(".intercept", ".sofa")
  )
```


 
 