---
title: Calculating Recency using dplyr
authors: 
  - admin
date: '2020-02-04'
categories:
  - Quick Solves
tags:
  - Quick Solves
  - dplyr
  - Recency
image:
  caption: ""
  focal_point: "Smart"
summary: "The idea here is to leverage dplyr power to calculate Customer recency."
highlight: true
math: true
output:
  blogdown::html_page:
    toc: false
    number_sections: false
    toc_depth: 4
    fig_caption: yes
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_knit$set(eval.after = 'fig.cap', message = FALSE, warning=FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
```

## The Problem
 
We have an Intern working on his Thesis Project in our office. He needed to calculate Customer recency, meaning he needed to know the amount of months since the last time the Customer made a Purchase. This was quite intriguing to me because it needs to combine some windows scoped functions with group by and some other things.

This is the problem with the expected solution:

```{r}
(data <- tibble::tribble(
  ~Client_ID, ~Date_ID, ~Purchase_Amount, ~Recency,
           1,        1,             2344,        0,
           1,        2,                0,        1,
           1,        3,                0,        2,
           1,        4,             5676,        0,
           1,        5,             4587,        0,
           1,        6,                0,        1,
           1,        7,                0,        2,
           1,        8,                0,        3,
           2,        1,             2500,        0,
           2,        2,             2634,        0,
           2,        3,                0,        1,
           2,        4,                0,        2,
           2,        5,                0,        3,
           2,        6,             4578,        0,
           2,        7,             4562,        0,
           2,        8,                0,        1
  )
)
```

As you may see every time I have a Purchase the counter needs to restart at 0 and then start counting how many dates have passed since the last purchase. Aditionally the counter needs to restart for new Customers.

## The solution

In this particular case I will detail the different steps of the solution because it can be tricky to get.

First I will create an auxiliary variable called has_purchased and a date_group. These variables need to be created at the client level, in order to make this easier I will use Client_ID 1 for demonstration purposes:

```{r}
data %>%
  filter(Client_ID == 1) %>% 
  mutate(has_purchased = as.numeric(Purchase_Amount > 0),
         date_group = cumsum(has_purchased))
```

Noticed that using the cumsum() function we can create groups in which the recency needs to restart. Everytime we change the date_group recency needs to come back to 0.

Then we can calculate the row_number by group and substract 1 and that's it.

```{r}
data %>%
  filter(Client_ID == 1) %>%
  mutate(
    has_purchased = as.numeric(Purchase_Amount > 0),
    date_group = cumsum(has_purchased)
  ) %>%
  group_by(date_group) %>%
  mutate(calculated_recency = row_number() - 1)
```

Now, in order to make this computationally efficient and generalize this solution to all of the customers we need to apply this by Client_ID. The way in which we'll do this is by using the group_modify() function. 
This function works very similarly to purrr's maps but applied to grouped data. The final solution looks like this:

```{r}
data %>%
  #grouped by client
  group_by(Client_ID) %>%
  group_modify(
    #This is the same pipeline showed before but applied to element .x that represents each group
    ~ .x %>%
      mutate(
        has_purchased = as.numeric(Purchase_Amount > 0),
        date_group = cumsum(has_purchased)
      ) %>%
      group_by(date_group) %>%
      mutate(calculated_recency = row_number() - 1)
    
  ) %>%
   select(-has_purchased, -date_group)

```

You'll notice calculated_recency matches with the expected results proving our solution works and a complicated calculation can be easily done using some group_by statements. 

I think the beauty of this solution is that we only used vectorized functions without applying any loop to run through the data by Client and by date_group which normally would take 2 nested for loops.


