---
title: Why this is failing?
authors: 
  - admin
date: '2019-10-16'
categories:
  - Quick Solves
tags:
  - Quick Solves
  - tidylog
image:
  caption: ""
  focal_point: "Smart"
summary: "How to avoid easy errors by getting short summaries of the operations applied to data."
highlight: true
math: true
output:
  blogdown::html_page:
    toc: false
    number_sections: false
    toc_depth: 4
    fig_caption: yes
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_knit$set(eval.after = 'fig.cap', message = FALSE, warning=FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(rlang))
```


This is a pretty typical issue. Specially when you have dealing with data a long time you just stop seeing obvious things, and you just can´t find solution to inexistant problems. For instance:

```{r}
mtcars %>%
  filter(cyl < 4)
```

You want to get the rows having cyl less or equal to 4 and for quite a while you keep getting 0 results.
Obviously something is wrong with the code but you just can´t notice it.

## The Solution

Well tidylog can give you an idea. Just load tidylog and watch:

```{r, echo = FALSE}
suppressPackageStartupMessages(library(tidylog))
```

```{r}
#loading the package this way to avoid verbose messages
library(tidylog)
mtcars %>%
  filter(cyl < 4)
```
Tidylog produces short log messages for dplyr and tidyr operations that help you understand what is happening with the data. Here definitely filter is incorrect, not producing an error but removing the 100% of the data, that is not what I was looking for.

Everytime you build a pipeline, tidylog will tell what is happening:

```{r}
mtcars %>%
  filter(cyl > 4) %>%
  select(-disp) %>%
  mutate( overall = rowMeans(.)) %>%
  summarize_all( ~ mean(.))
```

