---
title: My Final Project at the ML Diploma (Part II)
authors: 
  - admin
date: '2019-09-24'
categories:
  - Tutorial
tags:
  - Machine Learning
  - Data Import (haven)
  - Data Cleaning
image:
  caption: ""
  focal_point: "Smart"
summary: "Second Part of the ML Diploma Project. This time I´ll be showing some EDA"
highlight: true
math: true
output:
  blogdown::html_page:
    toc: true
    number_sections: true
    toc_depth: 4
    fig_caption: yes
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_knit$set(eval.after = 'fig.cap')
load("data-II.R")
suppressPackageStartupMessages(library(tidyverse))
```


Last time we conducted a high level cleansing of the data. Now it´s time to understand what is going on in it. In order to do that we´ll use a lot ggplot to visualize the data.

# Checking Numerical Distribution

In order to do this I should pick Numerical Variables one by one and create a ggplot.
This ould actually be quite tedious, why not to use the power of the tidyverse?

We will combine select_if and walk 2 to create histograms for every of the 28 Numerical Variables.

> Notice that in order to make walk work silently I had to add a print function that will use .x (every column) to create a histogram labeling it with .y that is the actual name of the current .x.


```{r}
# Take the data
data %>%
  # I select only data that is numerical
  select_if(is.numeric) %>%
  # I use walk 2 where .x is every numerical column seleted by select_if and
  #.y are the names of .x that will be used to add the proper label.
  walk2(names(.), ~ print( data %>%
                             ggplot(aes(.x)) + geom_histogram() + labs(x = .y))) 
```

# Checking Categorical Variables

Something equivalent can be done with categorical variables to check how they are distributed with the following code:

```{r}
data %>%
  select_if(is.factor) %>%
  summary()

```

In case you want something more visual you could go with this:

```{r}
# Take the data
data %>%
  # I select only data that is numerical
  select_if(is.factor) %>%
  # I use walk 2 where .x is every numerical column seleted by select_if and
  #.y are the names of .x that will be used to add the proper label.
  walk2(names(.), ~ print( data %>%
                             ggplot(aes(.x)) + geom_bar() + labs(x = .y))) 
```


# Chi-Square Test

What about performing a Chi-Square test to check the relationship between the Response variable and the Categorical Variables.

Let´s create a NSE function to apply Chi-Square using purrr.

We´ll use the Categorical Object created in the previous part to be looped over the chi-square function.

```{r}
#Listing all of the Categorical Variables according to Metadata
categorical <- c("ATM", "Branch", "CC", "CD", "DDA", "DirDep", "HMOwn", "ILS", "IRA", "InArea", "Ins", "Inv", "LOC", "MM", "MTG", "Moved", "NSF", "Res", "SDB", "Sav")
```


```{r}
#Loading rlang
suppressPackageStartupMessages(library(rlang))
  
#since I want to use var as a Non Standard Evaluation Variable I need to pass that variable using the Curly-Curly Operator. That way I don´t need to quote variables and can go directly into dplyr functions such as select.
  chi_comparison <- function(var){
    
   pred <- data %>%
     select({{ var }})
   
   #Performs Chi-Square test and returns p.value
   return(tibble(p_val = chisq.test(pred, data$Ins)$p.value))
  }

(independent <- categorical %>%
    map_dfr(chi_comparison) %>%
    cbind(independent = categorical) %>%
    filter(p_val > 0.05) 
)
```

This results in 4 Variables returning a p-value grater than 0.05. This means this variables are independent to the Response Variables, so no relationship between them exist, hence they could be removed from the model to build.

# Conclusion

A quick EDA has been performed using ggplot2 combined with purrr and dplyr.
 * It can be seen that Age and CRSore have distribution fairly close to Normal.
 * Income is right skewed.
 * Most of the Numerical Variables are higly concentrated at lower values.
 On the categorical side:
 * Most of the categorical variables show severe problems with class imbalances.
 * HMOwn, ILS, MTG and Moved seem to have no relationship with the Response Variable.
 * The Response Variable Ins show some imbalances but nothing to severe to be treated in a special way.


More to come on this problem. Stay Tuned!!!